Experimental and numerical studies of the effects of resonant and random excitations of the proton beam in the LHC, with applications to the design of a pulsed hollow electron lens for active halo control

### 2018-03-19 resex.tex
Revised text of transverse damper observations.

For all $k^{\mathrm{th}}$~turn pulsing patterns tested experimentally,
the transverse damper does in general not change any of the
observables studied during the experiments, explicitly losses,
emittance and beam distribution suggesting that indeed it is not
damping any resonant excitation. On the other hand, the transverse
damper considerably decreases any changes of the above parameters in
case of a random excitation. As an example, the vertical emittance
growth for the 7th- and \tenthtp\ in V is compared with a random excitation in V in Fig.~\ref{fig:damp}. The same observation is also made in case of losses as shown in Fig.~\ref{fig:ranexploss} comparing the \seventhtp\ in H+V with a random excitation in H+V. For the \seventhtp\ the two curves for the transverse damper active and not active almost coincide while for the random excitation a significant difference is visible. Until present, it is unknown why the transverse damper (ADT) appears to be incapable of damping any resonant excitation as it fulfills all requirements to detect as well as damp the caused oscillation:
\begin{itemize}
	\item The closed orbit distortion caused by the resonant excitation during the experiments is large enough to be detectable with the ADT pickups.
	\item The ADT compares the position of each bunch with its position in the previous turn in order to detect an orbit distortion to be damped. Therefore an oscillation due to any $k^{\mathrm{th}}$~turn pulsing with $k>1$ would be noticed and damped.
	\item The ADT is capable of damping the oscillation of each bunch individually. The ADT should therefore be capable of applying the correct corrective kick for each group of bunches seeing the same excitation amplitude.
\end{itemize}
Further studies are therefore needed in order to obtain a deeper understanding of the interaction of the transverse damper with a resonant excitation.

>>>>

During \kthtp\ experiments, the action of the transverse feedback and
damping system (ADT) did not significantly change any of the
observables, namely losses, emittances, and beam distributions. On the
other hand, the ADT considerably reduced any changes of the above
parameters in case of random excitations.

An example is shown in Fig.~\ref{fig:ranexploss}, where losses from
\seventhtp\ in H+V are compared with those measured during random
excitations in H+V. For \seventhtp, the results with and without
damper almost coincide, whereas a significant reduction was observed
for random excitations. As another example, vertical emittances for
7th- and \tenthtp\ in V are compared to those obtained during vertical
random excitations in Fig.~\ref{fig:damp}.

The reason why the transverse damper appeared to be capable of damping
random but not resonant excitations is unknown. In principle, it
fulfills the requirements needed to detect and damp the resulting
oscillations:
%
\begin{itemize}
\item the closed-orbit distortion caused by the resonant excitations
  was large enough to be detectable by the ADT pickups;
\item in order to detect the orbit distortion, the system compares the
  position of each bunch with its position in the previous turn;
  therefore, an oscillation due to \kthtp, with $k > 1$, should be
  detected and damped;
\item the transverse damper acts on each bunch individually, so it
  should be capable of applying the appropriate corrective kick to
  each group of bunches subject to the same excitation amplitude.
\end{itemize}

We report these observations in an attempt to advance the
understanding of the interaction between resonant excitations and
damping systems. Further studies are obviously needed.
